% -*-latex-*-
% Document name: install.tex
% Creator: Rob MacLeod [macleod@vissgi.cvrti.utah.edu]
% Last update: Fri Sep 15 00:33:54 2000 by Rob MacLeod
%    - created
% Last update: Wed Sep 27 14:54:43 2000 by Rob MacLeod
%    - some fixes for bug reporting and spell checking
% Last update: Sun Sep 24 16:41:08 2000 by Rob MacLeod
%    - Version 5.0Beta release edition
% Last update: Mon Jul 23 13:28:30 2001 by Rob MacLeod
%    - release 5.2
% Last update: Fri Jun 21 13:32:48 2002 by Rob Macleod
%    - release 5.3
% Last update: Fri Jan 24 20:00:00 2004 by Bryan Worthen
%    - release 5.4
% Last update: Thu May 19 12:35:15 2005 by Rob Macleod
%    - for version 6.3
% Last update: Fri Aug 26 12:34:14 2005 by Bryan Worthen
%    - for version 6.4
% Last update: Fri Feb 16 08:32:27 2007 by Rob Macleod
%    - version 6.5
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Installation}

\subsection{System requirements}

\map{} is written in standard C/C++ and uses the OpenGL and GTK+ libraries, 
both choices made to ensure broad portability of the program.

Note: As of \map{} 6.4 or greater, GTK+ version 2.4 or greater is required
to run \map{}.  Please update your dependencies according the the 
Windows, Linux, or Mac sections below.

  \begin{description}
    \item [All platforms: ]  OpenGL now comes standard on most systems.  
          Instructions on how to install GTK+ are described in detail below
          based on which platform you are installing.
          described below\mbox{}\\
          \begin{center}
          \begin{tabular}{|l|l|} \hline
            \multicolumn{2}{|c|}{Requirements for all systems.  } \\ \hline
          OpenGL libraries (GL and GLU) & version 1.1 +
     \footnote{A plus sign, ``+'', indicates ``or better''}\\
            OpenGL/window interface library (GLX)& \\
            \htmladdnormallink{GTK+}
            {http://www.gtk.org} 
            libraries and dependencies & version 2.4+ \\ \hline
          \end{tabular}
          \end{center}

            
    \item [Linux (i386): ] \map{} requires the OpenGL library, which is
              available 
          as the mesa library at \htmladdnormallink{www.mesa3d.org}
          {http://www.mesa3d.org} for any Linux platform.  For better
          performance, graphics cards from companies such as nVidia 
          \htmladdnormallink{(www.nvidia.com)}
          {http://www.nvidia.com/} usually provide OpenGL libraries.  \\
          \begin{center}
          \begin{tabular}{|l|l|} \hline
            \multicolumn{2}{|c|}{Requirements} \\ \hline
              Operating System & kernel 2.2.x\\
              Architecture & i386 (+ maybe PPC)\\
              Applications Binary Interface  &  libc2.1 \\
              \hline
            \multicolumn{2}{|c|}{Recommendations} \\ \hline
            Window system & XFree86 version 4.0 +\\
            Hardware & 3D graphics card (nVidia, 3dfx, ati)\\
            & 128 MB main memory \\ \hline
          \end{tabular}
          \end{center}
          
    \item [Windows: ] \mbox{}\\
          \begin{center}
          \begin{tabular}{|l|l|} \hline
            \multicolumn{2}{|c|}{Requirements} \\ \hline
              Operating System & W2K/NT4.0/9x\\
              Architecture & i386\\
              Applications Binary Interface  &  win32 \\
              \hline
            \multicolumn{2}{|c|}{Recommendations} \\ \hline
            Hardware & 3D graphics card (nVidia, 3dfx, ati)\\
            & 128 MB main memory \\ \hline
          \end{tabular}
          \end{center}

    \item [Mac OS X: ] \mbox{}\\
          \begin{center}
          \begin{tabular}{|l|l|} \hline
            \multicolumn{2}{|c|}{Requirements} \\ \hline
              Operating System & Mac OS 10.4(Tiger)\\
              Architecture & PPC or Intel\\
              \hline
            \multicolumn{2}{|c|}{Recommendations} \\ \hline
            Hardware & 3D graphics card (nVidia, 3dfx, ati)\\
            & Intel processor \\
            & $ge$512 MB main memory \\ \hline
          \end{tabular}
          \end{center}
          
   \item [SGI: ] we have dropped our support for SGI.  This platform, while
     so central to the origins of \map{} and OpenGL, is just not a
     significant desktop system in this day and age.  

  \end{description}
  
\subsection{General Installation}

Unfortunately, with our move to GTK+ for window support, it is not as easy
as past versions, which required just the download of an executable.  We
hope (in vain, perhaps) to be able to do that again in the future, but for
now we will attempt to make installation as easy as we can.  Simplified
instructions will be in a README file which comes with each package, and
are also listed below:

To download the software, use this URL
\htmladdnormallink{\texttt{www.sci.utah.edu/software/map3d.html}}
{http://www.sci.utah.edu/software/map3d.html}, and click on the
``Download'' button.  You'll need to sign into the SCI archive.  For
each of the installation instructions below, you can grab those file
from this page.


To test the installation, use the test files that accompany this
distribution.  Each has some script files included that show how to call
and execute \map{}.

\input{linux_install}
\input{windows_install}
\input{mac_install}
%\input{sgi_install}
\input{source_install}

\subsection{Documentation}

This document should have reached you either as a pdf file or via the
\htmladdnormallink{\map{} web site}
{http://www.sci.utah.edu/software/map3d.html}.  If you would like more
copies or the latest version, go to the
same web site and look for the links under Documentation:\\

\htmladdnormallink{\texttt{www.sci.utah.edu/software/map3d.html}}
{http://www.sci.utah.edu/software/map3d.html}

\subsection{Bug reporting}

We want to hear your response to using \map{} and especially to learn about
any bugs you may find.  They may be features, rather than bugs, but if so,
we will be happy to hear your impressions.

To submit a bug report please send email to map3d@sci.utah.edu or point your
browser at
\htmladdnormallink{\texttt{software.sci.utah.edu/bugzilla}}
{http://software.sci.utah.edu/bugzilla/} (you will need to register your e-mail
address) with the following information:

\begin{enumerate}
  \item Type and version of the operating system and hardware you are using.
  \item Version of \map{}.
  \item Description of the bug/feature you encountered.
  \item Suggestions for fixing the bug or altering the program behavior.
\end{enumerate}


\subsection{How to reach us}

We have established an email address for \map{},
\htmladdnormallink{map3d@sci.utah.edu}{mailto:map3d@sci.utah.edu}, and
web pages within the website \htmladdnormallink{www.sci.utah.edu/ncrr}
{http://www.sci.utah.edu/ncrr} dedicated to \map{}.  There is also a
majordomo mailing list for \map{} users called
\texttt{map3d-users@sci.utah.edu}.  To subscribe to this list, send email to
\texttt{majordomo@sci.utah.edu} with the following in the message body\\
%
\begin{verbatim}
      subscribe map3d-users
\end{verbatim}

Please let us know how you use \map{} and how we can make it better for
your purposes.  We can only develop this program with continued
support and the best way to achieve this is to show that others use the
program and find it helpful.



%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "manual"
%%% End: 
