 % -*-latex-*-
% Document name: newfeatures.tex
% Creator: Rob MacLeod [macleod@vissgi.cvrti.utah.edu]
% Last update: September 2, 2000 by Rob MacLeod
%    - created
% Last update: Mon Jul 23 13:28:30 2001 by Rob MacLeod
%    - release 5.2
% Last update: Fri Mar 1 20:00:00 2002 by Bryan Worthen
%    - release 5.3
% Last update: Fri Jan 24 20:00:00 2003 by Bryan Worthen
%    - release 5.4
% Last update: Tue Mar 30 20:00:00 2004 by Bryan Worthen
%    - release 6.0
% Last update: Wed Jun 30 13:28:30 2004 by Bryan Worthen
%    - release 6.1
% Last update: Tue Oct 19 13:28:30 2004 by Bryan Worthen
%    - release 6.2
% Last update: Wed May 11 13:28:30 2005 by Bryan Worthen
%    - release 6.3
% Last update: Thu May 19 12:35:15 2005 by Rob Macleod
%    - for version 6.3
% Last update: Web Jan 31 12:35:15 2005 by Bryan Worthen
%    - for version 6.4
% Last update: Fri Feb 16 08:16:29 2007 by Rob Macleod
%    - for version 6.5
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{What's New?}
\label{sec:new-features} 

In this section, we highlight the latest additions to \map{} in the
(vain?) hope that people will read at least this much of the manual
and be able to quickly make use of the latest and greatest that the program
offers.

\subsection{Version~\version{}: March, 2007}

This is the sixth version of the ``new'' \map{} with a GTK-based GUI.  We
are way beyond the complete functionality of the old GL
based version in almost all aspects, although it is sobering to notice that
some things that were easy in GL are still too hard in OpenGL. 

Some of the specific additions that you should notice over previous versions
include:
%
\begin{description}
  \item [GUI improvements: ] Some improvements in the GUIs, particularly in the
    fiducial control and contour spacing control.  See 
    Section~\ref{sec:interactive}.
  \item[Fiducial support:] fiducials are time markers that we often add to
    the time signals that we record or compute and suppoort for them in the
    new verison is hugely improved.  The first piece is the support for
    storing fiducials in the MATLAB files (see
    Section~\ref{sec:matlabdata}) (and also the CVRTI container files,
    Section~\ref{sec:tsdfcfile}) and the second part is the display features.
    It is now possible to visualize fiducials superimposed on potentials or
    to view them as separate fiducial maps.  This may all sound a little
    arcane but if you need this support, you will be very happy to see it.
  \item [Saving Matlab Geometry: ] This version allows the user to save matlab 
    geometry files, single or multiple surfaces per file.
  \item [Displaying an image inline with the Geometry] \map{} allows an
    image to be drawn in the background of the geometry.  See
    Section~\ref{sec:bgimage} for more details.
  \item [Bug fixes: ] not that the previous version had any bugs, but we
    found a few(!) more things to fix.
\end{description}

\subsection{In the works (``vapourware'')}
\label{sec:vapourware}

A small sampler of things that are in the works:

\begin{itemize}
  \item Saving the frames into actual movies.
  \item Dynamic menus that indicate the current parameter selections
    (a bit more in progress).
  \item New display modes for the display of vectors.
  \item Fixing the bugs listed on the bugs page (see
    Section~\ref{sec:bugs}).
\end{itemize}



%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "manual"
%%% End: 
