% -*-latex-*-
% Document name: windows_install.tex
% Creator: Bryan Worthen [worthen@sci.utah.edu]
% Last update: Fri Mar 26 20:00:00 2004 by Bryan Worthen
%    - release 6.0
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Windows Installation Instructions}
\label{sec:windows-install}

The Windows installation is relatively straightforward.  You'll need to 
download \map{}'s dependencies, then download \map{} itself.

\subsubsection{Windows Dependencies}

Download the \texttt{map3d-win-environment.zip} file from the \map{}
\htmladdnormallink{\texttt{download page}}
{http://www.sci.utah.edu/software/map3d.html} and unzip (using winzip,
native windows xp zip file browser, or another windows zip program) it into a location
of your choice - we will call that \texttt{INSTALL-DIR.}  It will create a
directory called \verb|INSTALL-DIR\map3d|.

Add \verb|INSTALL-DIR\map3d\lib| to your path.  To do this, open the
Control Panel, select System, and click on the `Advanced' tab on the top of
the screen.  Click on the Environment Variables button.  You should see a
variable called Path or PATH in the System Variables section.  Click on it,
and select Edit.  Go to the end of the line, add a semi-colon (;) and type
\verb|INSTALL-DIR\map3d\lib|.

\subsubsection{Windows Executable}

There is an executable of map3d in the environment directory.  We have also
placed an executable here to facilitate future downloads, so you only have
to download the environment once.  If you wish, download the
\texttt{map3d-6.4-windows.zip} file from the \map{} download page and unzip
it to a directory of your choice.  We will call that \texttt{RUN-DIR.} This
is the directory from which you will run \map{}.



%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "manual"
%%% End: 
